\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-03-20 paper/main.tex
.

\begin{table}[!tpb]
	\begin{tabular}{|p{2.2cm}|p{0.9cm}|p{0.75cm}|p{0.9cm}|p{0.75cm}|}
		\hline
		(a) STITCH results&\multicolumn{4}{c|}{PPI graph}\\
		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
		&Macro AUC&Micro $AUC_p$&Macro AUC&Micro $AUC_p$\\
		\hline
		MolPred&$0.69$&$0.65$&$0.69$&$0.67$\\
		\hline
		OntoPred&$0.88$&$0.87$&$0.92$&$0.93$\\
		\hline
		\name (MolPred + OntoPred) & $0.89$ & $0.90$&$\mathbf{0.93}$&$\mathbf{0.94}$\\
		\hline
		\hline
		(b) Yamanishi results&\multicolumn{4}{c|}{PPI graph}\\
		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
		&Macro AUC&Micro AUC&Macro AUC&Micro $AUC_p$\\
		\hline
		\name (MolPred + OntoPred) & $0.83$ & $0.82$&$0.84$&$0.84$\\
		\hline
	\end{tabular} \medskip

@@

	\caption{(a) + (b) Results for \name on STITCH and Yamanishi dataset evaluated with a 5-fold cross-validation. We hereby denote the molecular feature based predictor with \textit{MolPred}, while abbreviating the ontology based, top-down predictor with \textit{OntoPred}. (c) + (d) Results for various state of the art (c) drug--target interaction prediction methods on Yamanishi dataset and (d) drug--target affinity prediction methods on BIOSNAP dataset, evaluated on their original and the protein cross-validation splitting scheme, approximately reproducing the results of MolTrans \textbf{CITATION}.}
	\label{tab:Results}
\end{table}

@@

@@

@@

%\section{Findings}

\subsection{Deification of our method}
\begin{itemize}
	\item we built protein function and ontology based features based on DL2vec
	\item Ontology derived protein function focused features are highly predictive for dtis
	\item We built a versatile template for various features to test localization on the PPI graph
	\item normal GCNs don't work on PPI graph, as it is highly connected $\rightarrow$ needs stronger more expressive aggregation function $\rightarrow$ GENConv in residual blocks for better numerical stability
	\item protein functions localize on the PPI graph, while molecular features don't
	\item all AUROC in \% AUROC score on STITCH

@@

\subsection{How to insult other methods}
\begin{itemize}
	\item Only few other methods perform their split over proteins \citep{Survey2018}, DTI-CDF does it
	\item Running split over proteins is harder than, drug and drug protein pair split (see below table)
	\item this applies for both DTI prediction and drug target affinity prediction (and Saras gene-disease association)
	\item using indications is like cheating, as not applicable for searching new drugs
	\item drug indications are highly predictive for downstream tasks, but lack capability to differentiate highly related drugs/proteins
	\item Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)
	\item microAUC is a superior and more intuitive metric for drug repurposing $\rightarrow$ why for each protein and not for each drug

>>>>

\begin{table*}[!tpb]
  \centering
	\begin{tabular}{|p{2.2cm}|p{0.9cm}|p{0.75cm}|p{0.9cm}|p{0.75cm}|}
		\hline
		(a) STITCH results&\multicolumn{4}{c|}{PPI graph}\\
		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
		&Macro AUC&Micro $AUC_p$&Macro AUC&Micro $AUC_p$\\
		\hline
		MolPred&$0.69$&$0.65$&$0.69$&$0.67$\\
		\hline
		OntoPred&$0.88$&$0.87$&$0.92$&$0.93$\\
		\hline
		\name (MolPred + OntoPred) & $0.89$ & $0.90$&$\mathbf{0.93}$&$\mathbf{0.94}$\\
		\hline
		\hline
		(b) Yamanishi results&\multicolumn{4}{c|}{PPI graph}\\
		&\multicolumn{2}{c|}{without}&\multicolumn{2}{c|}{with}\\
		&Macro AUC&Micro AUC&Macro AUC&Micro $AUC_p$\\
		\hline
		\name (MolPred + OntoPred) & $0.83$ & $0.82$&$0.84$&$0.84$\\
		\hline
	\end{tabular} \medskip

@@

	\caption{(a) + (b) Results for \name on STITCH and Yamanishi dataset evaluated with a 5-fold cross-validation. We hereby denote the molecular feature based predictor with \textit{MolPred}, while abbreviating the ontology based, top-down predictor with \textit{OntoPred}. (c) + (d) Results for various state of the art (c) drug--target interaction prediction methods on Yamanishi dataset and (d) drug--target affinity prediction methods on BIOSNAP dataset, evaluated on their original and the protein cross-validation splitting scheme, approximately reproducing the results of MolTrans \textbf{CITATION}.}
	\label{tab:Results}
\end{table*}

@@

\subsection{``Bottom-up'' and ``top-down'' prediction of drug-target interactions}

What features localize on graph? Molecular interactions based on
structure not likely -- only possible to find binding sites to other
proteins.

Phenotype-based features localize...

@@

using indications is like cheating, as not applicable for searching new drugs

drug indications are highly predictive for downstream tasks, but lack capability to differentiate highly related drugs/proteins

we built protein function and ontology based features based on DL2vec

Ontology derived protein function focused features are highly predictive for dtis

We built a versatile template for various features to test localization on the PPI graph

normal GCNs don't work on PPI graph, as it is highly connected $\rightarrow$ needs stronger more expressive aggregation function $\rightarrow$ GENConv in residual blocks for better numerical stability

protein functions localize on the PPI graph, while molecular features don't

molecular features work for all (most?) drugs, top-down features only
for those that have a side-effect profile

\subsection{Evaluating drug--target interaction predictions}

Only few other methods perform their split over proteins \citep{Survey2018}, DTI-CDF does it

Running split over proteins is harder than, drug and drug protein pair split (see below table)

this applies for both DTI prediction and drug target affinity prediction (and Saras gene-disease association)

Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)

microAUC is a superior and more intuitive metric for drug repurposing $\rightarrow$ why for each protein and not for each drug

@@

A naive predictor (ranking proteins) and predict each drug similarly
achieves cutting edge performance ($87.5$ AUROC for whole dataset,
$85.5$ for 5-fold cross validation in drug-split) $\rightarrow$ No
prot focused microAUC possible. $\rightarrow$ hub proteins

Yamanishi Dataset is only partially suitable for comparing results, if
everybody just derives a suitable subset (DTIGEMS)

This also applies to drug target affinity prediction. We were hereby
able to roughly reproduce the results from MolTrans (Bioinformatics)
on BioSnap

%\subsection{Features to use in predicting drug targets}

\subsection{Conclusions}

Evaluating DTI predictions is non-trivial due to numerous biases in
the training and evaluation datasets.

Combining bottom-up and top-down features improves state of the art.

\subsection{Deification of our method}
\begin{itemize}
	\item all AUROC in \% AUROC score on STITCH

@@

\subsection{How to insult other methods}
\begin{itemize}
	\item Stratified Cross validation is suitable for training, but \textbf{NOT} for validating and testing (Uselessly high AUPRC)
	\item microAUC is a superior and more intuitive metric for drug repurposing $\rightarrow$ why for each protein and not for each drug
